Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-02-19 WholePlotTrajectories.Rmd
Manuscript update

```{r Options, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(cache=T, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '40%',fig.align="center")
# Installation des packages si nécessaire et chargement
Library <- function(Packages) {
  InstallAndLoad <- function(Package) {
    if (!Package %in% installed.packages()[, 1]) {install.packages(Package)}
    require(Package, character.only = TRUE)
  }
  invisible(sapply(Packages, InstallAndLoad))
}
# Ajouter les packages nécessaires ici
source("Scripts/DrawingTools.R")
#source("Scripts/GithubPages.R")
```

@@

Disturbance is defined as rapid changes in the amount and allocation of resources such as variation in the availability of light, the fluxes of heat and water, etc. 
It influences the environment and competition among species and combineds with environmental filters that select the pool of species fitting the environmental context to shape the composition and structure of communities in the long term: understanding ecosystems requires understanding their disturbance history [@White2001]. 
Although those changes proved impacting the proportion and identity of species, negligible or even positive impacts were highlighted when considering the community as a whole, in accordance with the intermediate disturbance hypothesis (IDH). 
The IDH that predicts a maximized species diversity at intermediate level of disturbance, however, mainly rely on the study of species richness which gives limited or even misleading information on forests recovery and functioning.
The mechanisms under which disturbance shapes communities diversity and structure involve the foster of certain species either stochastically, through the recruitment and dispersal limitation [@Hubbell1999] or determinisitcally through the selection of suited ecological types.
Disturbance indeed in the context of old growth, stable forests where competitive exclusion favors best competitor species causes a shift toward disturbance resistant species with rapid growth and efficient resources acquisition [@Haddad2008].
This is illustrated by a shift from slow growth with scarce resources and high competitive ability, "conservative strategies", to fast growth with prolific resourcesn, "acquisitive strategies".
A way to disentangle the mechanisms of disturbance dynamics would then to identify the shifts in communities functional types definde by the acquisitive or conservative functional traits and by the reproductive strategy through seed and early stage traits [@TerSteege2001;@Reich2014;@Herault2011].
Relevant functional types would be defined with the combination of resource acquisition traits: in that respect leaf and stem traits proved relevant in a large litterature [@Chave2009b;@Wright2004],and life history traits revealing the growth and reproductive strategy [@Westoby1998;@Herault2011]
Beyond diversity indices it is essential to analyse changes in taxonomic composition so that the impact on species per se and their particular impact on ecosystem can be followed [@╔Lavorel2002;@Bellwood2006;@Baraloto2012a] along with the ecological rules shaping communities assembly and abundance distribution [@Magurran1988].

@@

### Study site 
Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest corresponding to a tropical wet climate with mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) with a 3-months dry season (< 100 mm.months^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011]. 
Elevation ranges between 5 and 50 m and mean annual temperature is 26°C. 
Soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
The experiment corresponds to a network of twelve 6.25ha plots that have undergone a gradient of three logging, thinning and fuelwood treatments.
Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots. 
The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40 cm DBH for treatment 3 (T3). 
Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Blanc2009].

Table 1: Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.
---------------------------------------------------------------------------------------------------------------------------------
                          ***Timber***                 ***Thinning***              ***Fuelwood***           **%AGB lost**

-----------------    ------------------------    ------------------------    ------------------------    ------------------------
    ***Control***                                                                                                   0

-----------------    ------------------------    ------------------------    ------------------------    ------------------------
         ***T1***     DBH ≥ 50 cm,                                                                                               
                      commercial species,                                                                      [12%-33%]         
                      ~ 10 trees/ha

-----------------    ------------------------    ------------------------    ------------------------    ------------------------
         ***T2***     DBH ≥ 50 cm,                DBH ≥ 40 cm,                                                                   
                      commercial species,         non-valuable species,                                        [33%-56%]         
                      ~ 10 trees/ha               ~ 30 trees/ha

-----------------    ------------------------    ------------------------    ------------------------    ------------------------
         ***T3***     DBH ≥ 50 cm,                DBH ≥ 50 cm,                 40 cm ≤ DBH ≤ 50 cm                               
                      commercial species,         non-valuable species,        non-valuable species,           [33%-56%]         
                      ~ 10 trees/ha               ~ 15 trees/ha                ~ 15 trees/ha

---------------------------------------------------------------------------------------------------------------------------------

### Inventories protocol and dataset correction
The study site corresponds to a tropical rain forest with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae. 
In the twelve experimental plots of the experiment, all trees above 10 cm DBH are mapped and measured annually since 1984. 
During the inventories, trees are first identified with a vernacular name, assigned by the field team, and then with a scientific name, assigned by a botanist during regular botanical campaigns. 
In 1984, specific vernacular names were given to 62 commercial or common species whereas two identifiers separating palm trees and trees were given to all other species. 
The botanical campaigns to identify all trees at the species level started in 2003 and continued every 5 to 6 years.
This inventory protocol rasied some mehodological issues given the variability of botanical identification practices along time and among plots. 
Because vernacular name attributed to trees usually correspond to a variety of botanical species, some taxonomic uncertainty remains whenever a trustworthy botanical name cannot be given thereafter. 
This taxonomic uncertainty was accounted for in the composition and diversity metrics throught a Bayesian framework estimating the expectancy and variance of composition and diversity metrics. 
The uncertainty propagation is done by the extrapolation of complete inventories at genus level from real incomplete inventories, on the basis vernacular names replacement. The replacement of vernacular names uses a multinomial distribution M~v~([*s~1~, s~2~, …, s~N~*] ,[*α~1~, α~2~,…, α~N~*]) based on the observed association probability [*α~1~, α~2~,…, α~N~*] between each vernacular name *v* and the species [*s~1~, s~2~, …, s~N~*] recorded in the inventory.
See appendix 1 and @Aubry-Kientz2013 for the detailed methodology.

The functional metrics used a set of 6 functional traits values representing traits of the leaf and wood economics and life history traits.
The trait database encompassed 4 traits of the leaves: leaves thickness (*L_thick*), toughness (*L_tough*), total chlorophyll content (*LCC*) and specific leaf area (leaf area per unit dry mass; *SLA*), 2 traits of the stem: wood specific gravity (*SWG*) and bark thickness (*B_thick)* and two life-history traits: maximum specific height (*Hmax*) and the seed mass (*Smass*). 
Traits database came from the BRIDGE project (http://www.ecofog.gf/Bridge/index.html) where traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou.
To account for intraspecific traits variability, trait values at species level were assessed in sampling one of the measured individuals for iteration rather than averaging at species level. 
Missing trait values were filled using multivariate imputation by chained equation (mice), restricted to the samples pertaining to the next higher taxonomic level, in order to account for the phylogenetic signal of the functional traits. 
The dataset comprised 257 botanical species. Whenever a species inventoried was not in the dataset, it was attributed a set of traits values randomly sampled among species of the same next higher taxonomic level. 
All composition and diversity metrics correspond to the expectancy obtained after 50 iterations of the taxonomic uncertainty propagation framework and the missing trait values filling process.

@@

*Composition trajectories* _ Trajectories of taxonomic and functional composition after disturbance were examined through the ordination of successive inventories from 1989 (5 years after disturbance) in a two-dimensional space based on NMDS analysis. 
Classifications were based on either abundance-based Bray-Curtis \autoref{Fig1} or incidence-based Jaccard dissimilarity (data not shown).

Both taxonomic and functional composition were substantially affected by disturbance and positively correlated with its intensity.
*The distance from the initial condition was significantly time-dependent (P<0.01).* 
The taxonomic composition remained significantly different from the 1989 reference inventory until 30 years for all disturbed plots \autoref{Fig1}. 
From 20 years after disturbance, though, the distancing decreased which suggested a return toward the reference 1989 inventory and its resilience to disturbance \autoref{Fig1},\autoref{Fig2}. 
The functional composition distancing from the reference inventory was also positively correlated with disturbance intensity and similarly stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance \autoref{Fig2}. 
*The coordinates of the functional traits were measured in the two-dimensional ordination space that mapped plots evolution along time (graphs not showed) which revealed that the trajectories of disturbed plots along time headed toward acquisitive functional strategies (from high SWG to high SLA and chlorophyll content).*

@@

*Traits community weighted means* _ The trajectories of the community weighted means (CWM) of the 8 functional traits were measured at plot level \autoref{Fig3}.
**To be coming**

@@

```
### Disturbance impact on diversity

*Taxonomic diversity* _ Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined in relation to the 1994 inventories \autoref{Fig5}. 
*For the three indices the OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and a covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01).*

@@

The evolution along time of the difference of Richness, Shannon and Simpson diversity compared to reference inventories in 1989 remained even for 30 years after disturbance for the control plots whereas disturbed plots showed significant increase, positively correlated with the disturbance intensity \autoref{Fig5}. 
The richness stabilized for all disturbed plots from 20 years after disturbance \autoref{Fig5}. 
The Shannon and Simpson diversity also stabilized 20 years after disturbance for T1 and T2 plots \autoref{Fig5} whereas for 2 out of 3 plots T3 the Shannon and Simpson diversity continued increasing until 30 years after disturbance, illustrating a significant increase in communities’ evenness.

*Functional  diversity* _ The evolution along time of the Rao functinal diversity compared to reference inventories in 1989 remained even all along the survey for the control plots. 
All disturbed plots however showed an increase of the functional diversity that was positively correlated to the disturbance intensity \autoref{Fig6}. 30 after disturbance the diversity of disturbed plots remained significantly higher than this of control plots, thus whenever the initial disturbance intensity.
The slope of the initial diversity increase and the maximum diversity were both positively correlated with the disturbance intensity. 
*The OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and the covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01)*.
The trajectories of plot 7 from the T1 disturbance treatment was discarded because of too high taxonomic uncertainty preventing correct analysis.

@@

Our study underlined the significant but contrasting impact of disturbance on tropical forests taxonomic and functional diversity.
Specifically we unveiled an important decoupling between taxonomic evenness and functional characteristic of the communities, and by extension ecosystem functioning.
This highlighted the functional overlap between species in tropical forests,typical of their huge biodiversity, and intricately questioned the different facets of communities resilience.

In line with the assumtions of the IDH our 30 years survey confirmed the significant increase of communities evenness and functional richness after disturbance.
As it was demonstrated for the conditions 10 years after disturbance [@Molino2001] the directional changes in communities composition reflected environmental filtering after disturbance which would support the IDH as demonstrated in [@Molino2001] rather than pure stochastic dynamics as suggested in [@Hubbell1998].
The trajectories shaping a unimodal response to disturbance suggested the recovery of communities initial state as it was already suggested by the previous studies from the same site 10 years [@Molino2001] and 20 years [@Baraloto2012a] after disturbance.
Only one plot displayed a still increasing Simpson diversity, and all plots showed a stabilization or a decrease of the composition distancing from 1989 inventories. 
This recovery corroborate previous post-disturbance dynamics observed in a Bornean rainforest [@Cannon1998] where at large scale the richness was preserved although it was diminished locally or in Solomon islands where stands demonstrated a fast recovery and identical composition [@Burselm2000].

Although recovery seemed on its way the trajectories revealed significant changes in the ecological rules shaping communities over the 30 years of survey.
Disturbance is thought to shift exclusive competition favoring competitive species to environmental filtering selecting acquisitive and disturbance resistant species [@TerSteege2001;@Flores2006;@Haddad2008]
The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
In particular this is demonstrated by the evolution of seed mass classes proportions revealing a clear increase in small seeded and then large dispersive species.
Faster than for taxonomic composition and diversity though, functional characteristic returned to their initial values and the values observed for control plots.
This suggested some functional overlap in the pre-disturbance condition (*i.e* the usual tropical forests conditions). 
Such overlap and maybe functional versatility was already observed in highly diverse ecosystems [@Bellwood2006] and sugegsted as one mechanisms for high diversity maintenance.
Decoupled taxonomic and functional trajectories questions to what extent communities recovery is completed: ecosystem functioning might be quickly recoverd, and taxonomic composition following suit although along for much longer, but the resilience of the community itself might be altered as functional overlap themselves would be lost.
Along with probably other environmental modifications (like soils nutrient cycling and compaction [@Olander2005]) with alteration of communities resilience makes us expect different trajectories after subsequent disturbance.
As disturbance seemed to update communities toward more disturbance resistant ones [@Haddad2008] it is possible that eventually the disturbance regie observed has set another permanent condition, which would only be more diverse if good competitors, poor disturbance resitant survive.
> (Flores, 2006  for differential sapling success after disturbance)

@@

-	To go beyond whole community diversity and highlight the ecological mechanisms behind communities assembly and repond to the question of their resilience we should focus on recruitment.
It would also refine the role of environmental filters shaping communities assembly.

>>>>

```{r Options, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(cache=F, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '40%',fig.align="center")
# Installation des packages si nécessaire et chargement
Library <- function(Packages) {
  InstallAndLoad <- function(Package) {
    if (!Package %in% installed.packages()[, 1]) {install.packages(Package)}
    require(Package, character.only = TRUE)
  }
  invisible(sapply(Packages, InstallAndLoad))
}
# Ajouter les packages nécessaires ici
source("Scripts/DrawingTools.R")
#source("Scripts/GithubPages.R")
```

@@

Disturbance is defined as rapid changes in the amount and allocation of resources such as variation in the availability of light, the fluxes of heat and water, etc. 
It influences the environment and competition among species and combineds with environmental filters that select the pool of species fitting the environmental context to shape the composition and structure of communities in the long term: understanding ecosystems requires understanding their disturbance history [@White2001]. 
Although those changes proved impacting the proportion and identity of species, negligible or even positive impacts were highlighted when considering the community as a whole, in accordance with the intermediate disturbance hypothesis (IDH). 
The IDH that predicts a maximized species diversity at intermediate level of disturbance, however, mainly rely on the study of species richness which gives limited or even misleading information on forests recovery and functioning.
The mechanisms under which disturbance shapes communities diversity and structure involve the foster of certain species either stochastically, through the recruitment and dispersal limitation [@Hubbell1999] or determinisitcally through the selection of suited ecological types.
Disturbance indeed in the context of old growth, stable forests where competitive exclusion favors best competitor species causes a shift toward disturbance resistant species with rapid growth and efficient resources acquisition [@Haddad2008].
This is illustrated by a shift from slow growth with scarce resources and high competitive ability, "conservative strategies", to fast growth with prolific resourcesn, "acquisitive strategies".
A way to disentangle the mechanisms of disturbance dynamics would then to identify the shifts in communities functional types definde by the acquisitive or conservative functional traits and by the reproductive strategy through seed and early stage traits [@TerSteege2001;@Reich2014;@Herault2011].
Relevant functional types would be defined with the combination of resource acquisition traits: in that respect leaf and stem traits proved relevant in a large litterature [@Chave2009b;@Wright2004],and life history traits revealing the growth and reproductive strategy [@Westoby1998;@Herault2011]
Beyond diversity indices it is essential to analyse changes in taxonomic composition so that the impact on species per se and their particular impact on ecosystem can be followed [@Lavorel2002;@Bellwood2006;@Baraloto2012a] along with the ecological rules shaping communities assembly and abundance distribution [@Magurran1988].
> See Garnier 2001a for the description of traits functional impact

Since the functional traits chosen showed low intraspecific compared to interspecific variability [@Garnier2001a;@Westoby2006a] it its valid to assess the functional trajectories of communities through the functional values observed or the species assembly.

@@

### Study site 
Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest corresponding to a tropical wet climate with mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) with a 3-months dry season (< 100 mm.months^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011]. 
Elevation ranges between 5 and 50 m and mean annual temperature is 26°C. 
Soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
The experiment corresponds to a network of twelve 6.25ha plots that have undergone a gradient of three logging, thinning and fuelwood treatments \@ref(tab:Tab1).
Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots. 
The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40 cm DBH for treatment 3 (T3). 
Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Blanc2009].

Table : (\#tab:Tab1) Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.

------------------------------------------------------------------------------------------------------------------------
                          ***Timber***              ***Thinning***          ***Fuelwood***           **%AGB lost**

------------------  ------------------------ ------------------------- ------------------------ ------------------------
    ***Control***                                                                                          0

         ***T1***    DBH ≥ 50 cm,                                                                                       
                     commercial species,                                                                [12%-33%]         
                     ~ 10 trees/ha

         ***T2***    DBH ≥ 50 cm,              DBH ≥ 40 cm,                                                                 
                     commercial species,       non-valuable species,                                    [33%-56%]         
                     ~ 10 trees/ha             ~ 30 trees/ha

         ***T3***    DBH ≥ 50 cm,              DBH ≥ 50 cm,               40 cm ≤ DBH ≤ 50 cm                               
                     commercial species,       non-valuable species,      non-valuable species,         [35%-56%]         
                     ~ 10 trees/ha             ~ 15 trees/ha              ~ 15 trees/ha                                     
------------------------------------------------------------------------------------------------------------------------

### Inventories protocol and dataset correction

The study site corresponds to a tropical rain forest with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae. 
In the twelve experimental plots of the experiment, all trees above 10 cm DBH are mapped and measured annually since 1984. 
During the inventories, trees are first identified with a vernacular name, assigned by the field team, and then with a scientific name, assigned by a botanist during regular botanical campaigns. 
In 1984, specific vernacular names were given to 62 commercial or common species whereas two identifiers separating palm trees and trees were given to all other species. 
The botanical campaigns to identify all trees at the species level started in 2003 and continued every 5 to 6 years.
This inventory protocol rasied some mehodological issues given the variability of botanical identification practices along time and among plots. 
Because vernacular name attributed to trees usually correspond to a variety of botanical species, some taxonomic uncertainty remains whenever a trustworthy botanical name cannot be given thereafter. 
This taxonomic uncertainty was accounted for in the composition and diversity metrics throught a Bayesian framework estimating the expectancy and variance of composition and diversity metrics. 
The uncertainty propagation is done by the extrapolation of complete inventories at genus level from real incomplete inventories, on the basis vernacular names replacement. The replacement of vernacular names uses a multinomial distribution M~v~([*s~1~, s~2~, …, s~N~*] ,[*α~1~, α~2~,…, α~N~*]) based on the observed association probability [*α~1~, α~2~,…, α~N~*] between each vernacular name *v* and the species [*s~1~, s~2~, …, s~N~*] recorded in the inventory.
See appendix 1 and @Aubry-Kientz2013 for the detailed methodology.

The functional metrics used a set of 6 functional traits values representing traits of the leaf and wood economics and life history traits.
The trait database encompassed 4 traits of the leaves: leaves thickness, toughness, total chlorophyll content and specific leaf area (leaf area per unit dry mass), 2 traits of the stem: wood specific gravity and bark thickness and two life-history traits: maximum specific height and the seed mass. 
Traits database came from the BRIDGE project (http://www.ecofog.gf/Bridge/index.html) where traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou.
To account for intraspecific traits variability, trait values at species level were assessed in sampling one of the measured individuals for iteration rather than averaging at species level. 
Missing trait values were filled using multivariate imputation by chained equation (mice), restricted to the samples pertaining to the next higher taxonomic level, in order to account for the phylogenetic signal of the functional traits. 
The dataset comprised 257 botanical species. Whenever a species inventoried was not in the dataset, it was attributed a set of traits values randomly sampled among species of the same next higher taxonomic level. 
All composition and diversity metrics correspond to the expectancy obtained after 50 iterations of the taxonomic uncertainty propagation framework and the missing trait values filling process.

@@

*Composition trajectories* _ Trajectories of taxonomic and functional composition after disturbance were examined through the ordination of successive inventories from 1989 (5 years after disturbance) in a two-dimensional space based on NMDS analysis. 
Classifications were based on either abundance-based Bray-Curtis \@ref(fig:Fig1) or incidence-based Jaccard dissimilarity (data not shown).

Both taxonomic and functional composition were substantially affected by disturbance and positively correlated with its intensity.
*The distance from the initial condition was significantly time-dependent (P<0.01).* 
The taxonomic composition remained significantly different from the 1989 reference inventory until 30 years for all disturbed plots \@ref(fig:Fig1). 
From 20 years after disturbance, though, the distancing decreased which suggested a return toward the reference 1989 inventory and its resilience to disturbance \@ref(fig:Fig1),\@ref(fig:Fig2). 
The functional composition distancing from the reference inventory was also positively correlated with disturbance intensity and similarly stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance \@ref(fig:Fig2). 
*The coordinates of the functional traits were measured in the two-dimensional ordination space that mapped plots evolution along time (graphs not showed) which revealed that the trajectories of disturbed plots along time headed toward acquisitive functional strategies (from high SWG to high SLA and chlorophyll content).*

@@

*Traits community weighted means* _ The trajectories of the community weighted means (CWM) of the 8 functional traits were measured at plot level \@ref(fig:Fig3).
**To be coming**

@@

### Disturbance impact on diversity

*Taxonomic diversity* _ Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined in relation to the 1994 inventories \@ref(fig:Fig5). 
*For the three indices the OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and a covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01).*

@@

The evolution along time of the difference of Richness, Shannon and Simpson diversity compared to reference inventories in 1989 remained even for 30 years after disturbance for the control plots whereas disturbed plots showed significant increase, positively correlated with the disturbance intensity \@ref(fig:Fig5). 
The richness stabilized for all disturbed plots from 20 years after disturbance \@ref(fig:Fig5). 
The Shannon and Simpson diversity also stabilized 20 years after disturbance for T1 and T2 plots \@ref(fig:Fig5) whereas for 2 out of 3 plots T3 the Shannon and Simpson diversity continued increasing until 30 years after disturbance, illustrating a significant increase in communities’ evenness.

*Functional  diversity* _ The evolution along time of the Rao functinal diversity compared to reference inventories in 1989 remained even all along the survey for the control plots. 
All disturbed plots however showed an increase of the functional diversity that was positively correlated to the disturbance intensity \@ref(fig:Fig6). 30 after disturbance the diversity of disturbed plots remained significantly higher than this of control plots, thus whenever the initial disturbance intensity.
The slope of the initial diversity increase and the maximum diversity were both positively correlated with the disturbance intensity. 
*The OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and the covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01)*.
The trajectories of plot 7 from the T1 disturbance treatment was discarded because of too high taxonomic uncertainty preventing correct analysis.

@@

Our study showed the significant impact of disturbance on tropical forests communities.
The subsequent diversity trajectories confirmed the intermediate disturbance hypothesis debated for tropical forests through their correlation with disturbance intensity.
Besides it revealed the contrasting response of taxonomic and functional characteristics, specifically the decoupling between communities taxonomic evenness and their functional diversity and dominant functional traits values.
The long-term disturbance trajectories observed highlighted the unachieved but consistent recovery of communities assembly for the lowest disturbance intensity but questioned it after higher disturbance.

In line with the "vegetation quantity effect" suggesting that communities functional characteristics depends on dominant species rather that subtle changes in biodiversity [@reftobefound]. This would result in a faster recovery of ecosystem processes rates than in communities ocmposition and diversity as it mas largely demonstrated for grasslands [@Tilman1997; @Mouillot2011] and recently proved for tropical forests [@Lohbeck2015].  
This highlighted the functional overlap between species in tropical forests,typical of their huge biodiversity, and intricately questioned the different facets of communities resilience.
Ecosystem processes graped through the set of functional traits retained here however do not assess communities resilience and future dynamics.

In line with the assumtions of the IDH the observed diversity trajectories confirmed the increase of communities evenness and functional richness after disturbance.
The consistency of the IDH has been debated for the higly diverse tropical forests, some observing the importance of environmental filters to shape communities assembly and others arguing for a genuine stochastic recovery [@Hubbell1999;@Molino2001].
Consistently to the trajectories drawn here, a previous study of the same experiment revealed the directional changes in communities composition reflecting environmental filtering after disturbance [@Molino2001].
Disturbance regimes brought a unimodal response depending on disturbance intensity as suggested by the outcome of 10 years and 20 years post-disturbance analysis[@Molino2001;@Baraloto2012a].
Consistently with dynamicspreviously observed in a Bornean rainforest [@Cannon1998] or in Solomon islands [@Burslem2000], the unimodal response of diversities to disturbance suggested the maintenance of diversity levels, although for one of the intensively disturbed plot the community evenness was still increasing \@ref(fig:Fig5). 
What contrasted in our study though is that taxonomic composition did clearly recover \@ref(fig:Fig3), which may imply a long-term, if not irreversible, change in communities after disturbance.

Regarding the functionl trajectories and dominnant trait values of disturbed communities \@ref(fig:Fig3), disturbance appeared to favor acquisitive strategies, *i.e* species with high SLA, low wood density, small seeds, etc [@Westoby1998;@Wright2004;@Reich2014].
The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
This means a shift from exclusive competition favoring competitive species before disturbance to environmental filtering selecting acquisitive and disturbance resistant species.
In particular, the evolution seed mass classes proportions \@ref(fig:Fig4) revealed a clear increase in small seeded, therefore large dispersive, species [@TerSteege2001;@Flores2006;@Haddad2008].

The diversity and composition functional trajectories showed a decoupling between functional and taxonomic after the most intense disturbance (treatments T2 and T3).
Plots functional characteristic returned to their initial values, similar to those of control plots, faster than their taxonomic counterparts \@ref(fig:Fig2;fig:Fig3;fig:Fig4).
As functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2006] the recovery of communities functional characteristics would go along the recovery of ecosystems processes although taxonomic composition remain altered.
Such decoupling suggested the consistency of functional overlap and maybe functional versatility, implying a variety of functioning from the same characteristic, in the pre-disturbance condition.
Functional overlap has been observed in highly diverse ecosystems [@Bellwood2006] and suggested as one mechanisms for high diversity maintenance.
Knowing the decoupling between taxonomic and functional trajectories, this functional overlap may not be recovering after disturbance and therefore radically modify communities resilience and resistance [@Trenbath1999;@Elmqvist2003;@Diaz2005]. 
Given the observed lack of composition recovery, in favor of more disturbance resistant species [@Haddad2008], the likely decrease in functional overlap, and the probable changes in different life like lianas or epiphytes [@Martin2013] and environmental conditions like soils nutrient cycling and compaction [@Olander2005], the entirety of communities recovery after intense disturbance is questioned.
In this case, post-disturbance communities might have the same functioning, as suggested by the observed functional recovery, but composition would be permanently altered in the case of most intense disturbance and species contingent to undisturbed forests in particular would be impaired.
This threatens the global biodiversity along with the risk to loose cornerstone species and trigger unexpected consequences [@Diaz2005;@Gardner2007].

To confirm this long-term communities changes and highlight the ecological rules shaping communities assembly, we suggest a focus on recruitment dynamics, that would follow the diversity and functional characteristic of recruited communities and, by extension, of the forest to be.

@@
